Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-08-17 doc/RMD/pop-diff.Rmd
Add automatic DAPC reporting in manuscript

```{r dapc-region, fig.width = 9, fig.height = 8, cache = TRUE}
scatter(region.dapc, 
        scree.pca = TRUE,
        bg = "grey95",
        col = "black",
        # col = rev(funky(nlevels(strata(dat11cc)$Region))), 
        pch = seq(nlevels(strata(dat11cc)$Region)),
        legend = "true",
        posi.leg = "topleft",
        posi.pca = "topright"
        )
ggcompoplot(region.dapc, setPop(dat11cc, ~Region), pal = rev(funky(nlevels(strata(dat11cc)$Region))), cols = 2)
```

>>>>

```{r dapc-region, fig.width = 9, fig.height = 8, cache = TRUE}
region.dapc
scatter(region.dapc, 
        scree.pca = TRUE,
        bg = "grey95",
        col = "black",
        # col = rev(funky(nlevels(strata(dat11cc)$Region))), 
        pch = seq(nlevels(strata(dat11cc)$Region)),
        legend = "true",
        posi.leg = "topleft",
        posi.pca = "topright"
        )
ggcompoplot(region.dapc, setPop(dat11cc, ~Region), pal = rev(funky(nlevels(strata(dat11cc)$Region))), cols = 2)
```

### 2017-08-17 doc/manuscript/manuscript.Rmd
Add automatic DAPC reporting in manuscript

@@

Discriminant Analysis of Principal Components (DAPC) was performed grouping by Region with the first 21 principal components, representing 88% of the total variance. 
The first discriminant axis (representing 64% of the discriminatory power) separated the centroid for the Mexico isolates from the rest of the data, indicating strong differentiation (Fig. \@ref(DAPC)b). 
The second discriminant axis separated the centroid for the California isolates.
The mean population assignment probabilities for all populations with n > 10 showed that only isolates from Mexico, CA, and France had > 50% probabilities of being reassigned to their source populations (Fig. \@ref(DAPC)a). 
DAPC grouping by cultivar used the first 20 principal components, representing 89% of the total variance.
The first two discriminant axes failed to separate any of the cultivars where the mean posterior assignment probabilities were 34.0% (G122), 36.0% (Beryl), and 30.1% (Bunsi).
DAPC grouping by Region and Year used the first 20 principal components, representing 87.7% of the total variance.
The Midwestern populations (NE, MI, CO, ND) did not appear to have any variation across time in contrast to WA, which showed a shift in population structure in the last year of sampling, 2008 (\@ref(DAPC-RY)).
Further analysis of this population revealed that all 12 isolates in WA circa 2008 originated in a wmn and nine of these genotypes were shared with CA, and three were shared with France (Fig. \@ref(community-graph), \@ref(community-graph16)).

>>>>

```{r initiate, echo = FALSE, message = FALSE, warning = FALSE}
library("poppr")
load("../../data/sclerotinia_16_loci.rda")
dat11cc <- clonecorrect(dat11, ~Region/Source/Host/Year)

dapc_region      <- dapc(setPop(dat11cc, ~Region), n.da = 13, n.pca = 20)
dapc_region_year <- dapc(setPop(dat11cc, ~Region/Year), n.da = 12, n.pca = 22)
dapc_host <- dat11cc %>% 
  setPop(~Source) %>%
  popsub("wmn") %>%
  setPop(~Host) %>% 
  popsub(c("Beryl", "Bunsi", "G122")) %>%
  dapc(n.da = 2, n.pca = 20)

prettify <- function(x, i) signif((1 - x[i])*100, 3)
paxis <- function(d, i = 1 ){
  res <- cumsum(d$eig)/sum(d$eig)
  prettify(res, i)
}
passp <- function(d, i = 1){
  res <- apply(d$posterior, 2, mean, na.rm = TRUE)
  prettify(1 - res, i)
}
```

@@

Discriminant Analysis of Principal Components (DAPC) was performed grouping by Region with the first `r dapc_region$n.pca` principal components, representing `r signif(dapc_region$var*100, 3)`% of the total variance. 
The first discriminant axis (representing `r paxis(dapc_region)`% of the discriminatory power) separated the centroid for the Mexico isolates from the rest of the data, indicating strong differentiation (Fig. \@ref(DAPC)b). 
The second discriminant axis, representing `r paxis(dapc_region, 2)`% of the discriminatory power, separated the centroid for the California isolates.
The mean population assignment probabilities for all populations with n > 10 showed that only isolates from Mexico, CA, and France had > 50% probabilities of being reassigned to their source populations (Fig. \@ref(DAPC)a). 
DAPC grouping by cultivar used the first 20 principal components, representing 89% of the total variance.
The first two discriminant axes failed to separate any of the cultivars where the mean posterior assignment probabilities were `r passp(dapc_host, "G122")`% (G122), `r passp(dapc_host, "Beryl")`% (Beryl), and `r passp(dapc_host, "Bunsi")`% (Bunsi).
DAPC grouping by Region and Year used the first `r dapc_region_year$n.pca` principal components, representing `r signif(dapc_region_year$var*100, 3)`% of the total variance.
The Midwestern populations (NE, MI, CO, ND) did not appear to have any variation across time in contrast to WA, which showed a shift in population structure in the last year of sampling, 2008 (Fig. \@ref(DAPC-RY)).
Further analysis of this population revealed that all 12 isolates in WA circa 2008 originated in a wmn and nine of these genotypes were shared with CA, and three were shared with France (Fig. \@ref(community-graph), \@ref(community-graph16)).
